---
output: github_document
bibliography: vignettes/resources/references.json
link-citations: true
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# daedalus: Optimise economic, social and health trade-offs in a pandemic

<!-- badges: start -->
[![Project Status: Concept – Minimal or no implementation has been done yet, or the repository is only intended to be a limited example, demo, or proof-of-concept.](https://www.repostatus.org/badges/latest/concept.svg)](https://www.repostatus.org/#concept)
[![R build status](https://github.com/j-idea/daedalus/workflows/R-CMD-check/badge.svg)](https://github.com/j-idea/daedalus/actions/workflows/R-CMD-check.yaml)
[![Codecov test coverage](https://codecov.io/gh/j-idea/daedalus/branch/main/graph/badge.svg)](https://app.codecov.io/gh/j-idea/daedalus?branch=main)
[![CRAN status](https://www.r-pkg.org/badges/version/daedalus)](https://CRAN.R-project.org/package=daedalus)
<!-- badges: end -->

_daedalus_ implements the integrated epidemiological and economic model in @haw2022.

## Installation

You can install the development version of daedalus from [GitHub](https://github.com/) using the _remotes_ package.

``` r
# install.packages("remotes")
remotes::install_github("j-idea/daedalus", upgrade = FALSE)
```

## Quick start

```{r libraries}
library(daedalus)
```

### Prepare contact matrix and initial conditions

```{r setup}
polymod <- socialmixr::polymod
# suppress messages that clog up test output
suppressMessages(
  contact_data <- socialmixr::contact_matrix(
    polymod,
    countries = "United Kingdom",
    age.limits = c(0, 5, 20, 65),
    symmetric = TRUE
  )
)

# get demography vector
demography <- contact_data[["demography"]][["population"]]

# prepare contact matrix
contact_matrix <- t(contact_data[["matrix"]]) / demography

# initial state: one in every 1 million is infected
initial_i <- 1e-6
initial_state <- c(
  S = 1.0 - initial_i, E = 0.0,
  Is = initial_i, Ia = 0.0,
  H = 0.0, R = 0.0, D = 0.0
)

# build for all age groups
initial_state <- rbind(
  initial_state,
  initial_state,
  initial_state,
  initial_state,
  deparse.level = 0L
)

# multiply by demography vector for absolute values
initial_state <- initial_state * demography
```

### Run model

The model runs for 100 timesteps (assumed to be days) by default.

```{r run_model}
output <- daedalus(
  initial_state,
  parameters = default_parameters(contact_matrix = contact_matrix)
)
```

Get the data in long or 'tidy' format using `prepare_output()`.

```{r get_data}
data <- prepare_output(output)

head(data)
```

## Related projects

WIP.

## References
