---
output: github_document
bibliography: vignettes/resources/references.json
link-citations: true
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# daedalus: Model health, social, and economic costs of a pandemic

<!-- badges: start -->
[![Project Status: Concept – Minimal or no implementation has been done yet, or the repository is only intended to be a limited example, demo, or proof-of-concept.](https://www.repostatus.org/badges/latest/concept.svg)](https://www.repostatus.org/#concept)
[![R build status](https://github.com/jameel-institute/daedalus/workflows/R-CMD-check/badge.svg)](https://github.com/jameel-institute/daedalus/actions/workflows/R-CMD-check.yaml)
[![Codecov test coverage](https://codecov.io/gh/jameel-institute/daedalus/branch/main/graph/badge.svg)](https://app.codecov.io/gh/jameel-institute/daedalus?branch=main)
[![CRAN status](https://www.r-pkg.org/badges/version/daedalus)](https://CRAN.R-project.org/package=daedalus)
<!-- badges: end -->

_daedalus_ implements the integrated epidemiological and economic model in @haw2022.

## Installation

You can install the development version of daedalus from this repository using the _pak_ package, or from the Jameel Institute R-universe.

```r
# install.packages("pak")
pak::pak("jameel-institute/daedalus", upgrade = FALSE)

# installation from R-universe
install.packages(
  "daedalus", 
  repos = c(
    "https://jameel-institute.r-universe.dev", "https://cloud.r-project.org"
  )
)
```

### Installation notes

_daedalus_ is under active development. To use a version of _daedalus_ associated with major milestones, install a specific version from the list below.

- IDM Thailand 2024: `pak::pak("jameel-institute/daedalus@v0.1.0")`

- IfG workshop 2025 (with real time modelling): `pak::pak("jameel-institute/daedalus@v0.2.0")`

### Known issues

We have found the following issues with released versions:

1. `v0.1.0`: Consumer-worker contacts are not correctly handled, leading to higher and faster epidemic peaks.

2. `v0.2.0`: Consumer-worker contacts are not included in the model, leading to lower and later epidemic peaks. The 'recovered' compartment does not include recoveries from hospitalisation. Please use the helper function `get_epidemic_summary()` to get epidemic sizes.

## Quick start

The model can be run for any country or territory in the `country_names` list by passing the country name to the function.
This automatically pulls country-specific demographic and economic data, which is included in the package, into the model (see the ['Get started' vignette](https://jameel-institute.github.io/daedalus/articles/daedalus.html) for more details).

```{r libraries}
library(daedalus)

# run model for Canada
data <- daedalus("Canada", "influenza_1918")

# get pandemic costs as a total in million dollars
get_costs(data, "total")

# disaggregate total for economic, education, and health costs
get_costs(data, "domain")
```

Users can select infection parameters from among seven epidemics caused by directly-transmitted viral respiratory pathogens, which are stored in the stand-alone helper package `daedalus.data`. These can be called as `daedalus.data::infection_data`, while epidemic identifiers are stored as `daedalus.data::epidemic_names`.

Users can override default country contact data and epidemic-specific infection arguments by passing custom classes to `daedalus()`; see the package website for more details.

Users can also model the implementation of pandemic response measures: for more on this see the documentation for the main model function `daedalus()`, and the vignette on modelling interventions on the package website.

## Related projects

_daedalus_ is an R implementation of the scenario model from a [project on the economics of pandemic preparedness](https://github.com/robj411/p2_drivers).

[_daedalus.data_](https://github.com/jameel-institute/daedalus.data) is a stand-alone helper R package to handle input data for the _daedalus_ model. Its purpose is to allow users to flexibly manipulate or change epidemiological and economic data, without needing to modify the model itself.

## References
