---
output: github_document
bibliography: vignettes/resources/references.json
link-citations: true
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# daedalus: Optimise economic, social and health trade-offs in a pandemic

<!-- badges: start -->
[![Project Status: Concept – Minimal or no implementation has been done yet, or the repository is only intended to be a limited example, demo, or proof-of-concept.](https://www.repostatus.org/badges/latest/concept.svg)](https://www.repostatus.org/#concept)
[![R build status](https://github.com/j-idea/daedalus/workflows/R-CMD-check/badge.svg)](https://github.com/j-idea/daedalus/actions/workflows/R-CMD-check.yaml)
[![Codecov test coverage](https://codecov.io/gh/j-idea/daedalus/branch/main/graph/badge.svg)](https://app.codecov.io/gh/j-idea/daedalus?branch=main)
[![CRAN status](https://www.r-pkg.org/badges/version/daedalus)](https://CRAN.R-project.org/package=daedalus)
<!-- badges: end -->

_daedalus_ implements the integrated epidemiological and economic model in @haw2022.

## Installation

You can install the development version of daedalus from [GitHub](https://github.com/) using the _remotes_ package.

``` r
# install.packages("remotes")
remotes::install_github("j-idea/daedalus", upgrade = FALSE)
```

## Quick start

The model can be run for any country or territory in the `country_names` list by passing the country name to the function.
This automatically pulls country-specific demographic and economic data, which is included in the package, into the model (see the ['Get started' vignette](https://jameel-institute.github.io/daedalus/articles/daedalus.html) for more details).

```{r libraries}
library(daedalus)

# run model for Canada
output <- daedalus("Canada", "influenza_1918")
```

Users can select infection parameters from among seven epidemics caused by directly-transmitted viral respiratory pathogens, which are stored in the package as `daedalus::infection_data`: SARS 2004 (SARS-CoV-1), influenza 2009 (influenza A H1N1), influenza 1957 (influenza A H2N2), influenza 1918 (influenza A H1N1), Covid-19 wild type (SARS-Cov-2 wild type), Covid-19 Omicron (SARS-CoV-2 omicron), and Covid-19 Delta (SARS-CoV-2 delta).

Epidemic identifiers are stored as `daedalus::epidemic_names`.

Users can override default country contact data and epidemic-specific infection arguments by passing named lists of parameters to `daedalus()`; see the function documentation for more details.

A helper function prepares the data in long or 'tidy' format using `prepare_output()`.

```{r get_data}
data <- prepare_output(output)

head(data)
```

## Related projects

_daedalus_ is an R implementation of the scenario model from a [project on the economics of pandemic preparedness](https://github.com/robj411/p2_drivers).

## References
