---
title: "Getting started with DAEDALUS"
output: rmarkdown::html_vignette
bibliography: resources/references.json
link-citations: true
vignette: >
  %\VignetteIndexEntry{Getting started with DAEDALUS}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This vignette shows how to get started with the DAEDALUS model adapted from @haw2022 in R.

```{r libraries}
# load the package
library(daedalus)
library(ggplot2)
```

The model can be run for any country or territory included in package data simply by passing its name to `daedalus()`.
The `country_names` vector holds a list of country and territory names for which data is available.

Countries and territories are represented by the `<country>` class, which can be created by calling `country()` on a country name. This collects package data on the country into a single object, making it easier to work with.

```{r make_country}
# make a <country> representing Canada (chosen for its short name)
country_canada <- country("Canada")

# print to examine; only some essential information is shown
country_canada
```

The package provides data from @walker2020 on country demography, country workforce per economic sector, and social contacts between age groups in `country_data`.
The package also provides data from @jarvis2024 on workplace contacts in economic sectors.
Both datasets are accessed by internal functions to reduce the need for user input.

```{r run_model}
# simulate a Covid-19 wild type outbreak
data <- daedalus(country_canada, "sars_cov_2_pre_alpha")
```

Users can quickly model one of seven epidemics by passing a name from `daedalus::epidemic_names` to the `epidemic` argument, which is the second argument.
The package data object `daedalus::epidemic_names` gives the allowed epidemic identifiers.

```{r epidemic_names}
daedalus::epidemic_names
```

The model runs for 300 timesteps by default; timesteps should be interpreted as days since model parameters are in terms of days.
The `daedalus()` function documentation has more information on the infection parameter defaults in the **Details** section.

Users can also override epidemic-specific infection parameter values as well as change the initial conditions by passing arguments to `daedalus()` as lists: `infect_params_manual` and `initial_state_manual`, which override the infection parameters and the initial state defaults respectively.
The function documentation for `daedalus::daedalus()` details which parameters are accepted.

```r
# not run
# e.g. change the transmission rate by changing r0
daedalus(
  country("United Kingdom"), 
  "influenza_1918", 
  infect_params_manual = list(r0 = 1.3)
)

# e.g. change the hospitalisation risk profile to model
# 1918 Influenza epidemic but with risk of severe disease
# increasing with age
daedalus(
  country("China"), "influenza_1918",
  infect_params_manual = list(eta = c(2.0, 3.0, 5.0, 15.0))
)
```

Plot the data to view the epidemic curve.

```{r plot_epidemic}
ggplot(
  data[data$compartment == "infect_symp" & data$age_group == "20-65", ]
) +
  geom_line(
    aes(time, value, colour = econ_sector),
    show.legend = FALSE
  ) +
  facet_wrap(
    facets = vars(age_group)
  )
```

## References
