---
title: "Daedalus features in development"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Daedalus features in development}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(daedalus)
library(ggplot2)
library(dplyr)
```

This vignette is intended as a visual check on _daedalus_ features in development.

## Basic `daedalus2()` usage

This example shows (checks) that `daedalus2()` output works with pre-built function `get_incidence()` written for `daedalus()`.

```{r}
output <- daedalus2("GBR", "influenza_2009")

# get incidence data using function developed for daedalus()
data <- get_incidence(output, "infections")

ggplot(data) +
  geom_line(
    aes(time, value)
  ) +
  labs(x = "Time", y = "New infections")
```

## Vaccination

This section aims to be a visual check on vaccination in `daedalus2()` as this function does not currently log new vaccinations.

```{r}
# high vaccination rate but low uptake limit
vax <- daedalus_vaccination("high", uptake_limit = 40, start_time = 30)
output <- daedalus2("GBR", "influenza_2009", vaccine_investment = vax)

data <- get_data(output) %>%
  filter(vaccine_group == "vaccinated", compartment == "susceptible") %>%
  group_by(time) %>%
  summarise(
    value = sum(value)
  )

# check that vaccinations begin at time = 100
ggplot(data, aes(time, value)) +
  geom_line() +
  labs(
    x = "Time", y = "Vaccinated susceptibles (prevalence)",
    caption = "Vaccinated susceptibles decrease due to infections."
  )
```

Check that fully protective (non-leaky) vaccination does not see a mid-epidemic reduction in vaccinated susceptibles.

```{r}
# efficacy is 100 for non-leaky vaccination
vax <- daedalus_vaccination(
  "high",
  start_time = 30, efficacy = 100,
  uptake_limit = 40
)
output <- daedalus2("GBR", "influenza_2009", vaccine_investment = vax)

data <- get_data(output) %>%
  filter(vaccine_group == "vaccinated", compartment == "susceptible") %>%
  group_by(time) %>%
  summarise(
    value = sum(value)
  )

# check that vaccinations begin at time = 100
ggplot(data, aes(time, value)) +
  geom_line() +
  labs(
    x = "Time", y = "Vaccinated susceptibles (prevalence)",
    caption = "Vaccinated susceptibles do not decrease due to infections."
  )
```

## Events

This section is a visual check on events.

```{r}
# trigger event on low hospital capacity
x <- daedalus_country("GBR")
x$hospital_capacity <- 1e3
x

output_npi <- daedalus2(
  x, "influenza_2009",
  response_strategy = "elimination"
)
output_no_npi <- daedalus2(
  x, "influenza_2009"
)

# get incidence data and check for differences
data_npi <- get_incidence(output_npi, "infections")
data_no_npi <- get_incidence(output_no_npi, "infections")

# add scenario marker and plot comparison
data_npi$scenario <- "npi"
data_no_npi$scenario <- "no_npi"

data <- bind_rows(data_npi, data_no_npi)

ggplot(data, aes(time, value, col = scenario)) +
  geom_line() +
  labs(
    x = "Time", y = "Incidence"
  )
```
