---
title: "Getting started with DAEDALUS"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with DAEDALUS}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This vignette shows how to get started with the DAEDALUS model in R.

```{r libraries}
# load the package
library(daedalus)
library(dplyr)
library(ggplot2)
```

### Prepare contact matrix and initial conditions

Load a contact matrix for the U.K. from the [_socialmixr_ package](https://cran.r-project.org/package=socialmixr).

```{r setup}
polymod <- socialmixr::polymod
# suppress messages that clog up test output
contact_data <- socialmixr::contact_matrix(
  polymod,
  countries = "United Kingdom",
  age.limits = c(0L, 5L, 20L, 65L),
  symmetric = TRUE
)

# get demography vector
demography <- contact_data[["demography"]][["population"]]

# prepare contact matrix
contact_matrix <- t(contact_data[["matrix"]]) / demography

# initial state: one in every 1 million is infected
initial_i <- 1e-6
initial_state <- c(
  S = 1.0 - initial_i, E = 0.0,
  Is = initial_i, Ia = 0.0,
  H = 0.0, R = 0.0, D = 0.0
)

# build for all age groups
initial_state <- rbind(
  initial_state,
  initial_state,
  initial_state,
  initial_state
)

# multiply by demography vector for absolute values
initial_state <- initial_state * demography
```

Prepare the initial state container, a 3D tensor.

Constants `N_AGE_GROUPS`, `N_EPI_COMPARTMENTS`, `N_ECON_SECTORS`, `i_NOT_WORKING` and `i_WORKING_AGE` are package constants.

```{r}
# construct state array, all working age individuals are uniformly
# distributed into economic sectors including non-working (hence + 1)
state <- array(
  0.0,
  dim = c(N_AGE_GROUPS, N_EPI_COMPARTMENTS, N_ECON_STRATA)
)
state[, , i_NOT_WORKING] <- initial_state
state[i_WORKING_AGE, , ] <- matrix(
  state[i_WORKING_AGE, , i_NOT_WORKING] * 1.0 / (N_ECON_STRATA),
  nrow = N_EPI_COMPARTMENTS, ncol = N_ECON_STRATA
)

# prepare workplace contacts
cw <- rep(5.0, N_ECON_SECTORS)

# prepare consumer worker contacts
consumer_contacts_per_sector <- withr::with_seed(
  0L, runif(N_ECON_SECTORS, min = 5.0, max = 10.0)
)
cmcw <- matrix(
  consumer_contacts_per_sector,
  nrow = N_ECON_SECTORS, ncol = N_AGE_GROUPS
) * demography / sum(demography)
```

### Run model

The model runs for 100 timesteps (assumed to be days) by default.

```{r run_model}
output <- daedalus(
  state,
  parameters = default_parameters(
    contact_matrix = contact_matrix,
    contacts_workplace = cw,
    contacts_consumer_worker = cmcw
  )
)
```

Get the data in long or 'tidy' format using `prepare_output()`.

```{r get_data}
data <- prepare_output(output)

head(data)
```

Plot the data to view the epidemic curve.

```{r plot_epidemic}
data |>
  filter(compartment == "infect_symp") |>
  ggplot() +
  geom_line(
    aes(time, value, colour = econ_sector),
    show.legend = FALSE,
    position = "jitter"
  ) +
  facet_wrap(
    facets = vars(age_group)
  )
```
