---
title: "Value of a statistical life"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Value of a statistical life}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This vignette explains how the value of a statistical life (VSL, also referred to hereafter as 'life-value') is calculated in _daedalus_.

The methodology used here has been chosen so as to be generalisable for the widest possible range of countries and territories, while using publicly available data.

This approach balances the ability to run _daedalus_ scenarios for a large number of countries, and the effort required to obtain and update official life-values from national governments.

::: {.alert .alert-info}
**Note that** the VSL values used in a call to `daedalus()` for any country can always be substituted with values considered to be more appropriate (e.g., from a national health ministry).
:::

## Value of a life year

The value of a life year (VLY) in a country is assumed to be the gross national income (GNI) per capita scaled for purchasing power parity in international dollars.

Country GNI values are for the most recent year available, and are taken from the [World Bank's World Development Indicators](https://databank.worldbank.org/source/world-development-indicators).

::: {.alert .alert-info}
**Note that** there is no discounting applied to older age groups, that is, all remaining years of life are valued the same, and are not adjusted for disabilities or quality.
:::

## Life expectancy

Life-expectancy data in terms of years for each country and territory is initially accessed for age-bins and separated by sex.

The life-expectancy data is then aggregated by the DAEDALUS age bins: 0 -- 4, 5 -- 19 (school age), 20 -- 65 (working age), 65+ (retired or not working).

::: {.alert .alert-warning}
**Note that** the aggregation of is a simple mean across all age-bins in the raw data; a more appropriate approach which we aim to add in future would be a weighted mean that takes into account the size of each age-bin in the raw data.
:::

This results in four life-expectancy values for each country, each corresponding to one of the four DAEDALUS age groups.

## Value of a statistical life

The value of a statistical life for an individual of country $c$ in age group $i$ denoted as $\text{VSL}_{c_i}$ is the product of the country-specific VLY ($\text{VLY}_{c}$) and the country- and age-specific life expectancy ($T_{c_i}$), and depends on the country and age-group of the individual; this can be expressed as:

$$\text{VSL}_{c_i} = \text{VLY}_c \times T_{c_i}$$

Essentially, the VSL lost due to a death is the country-specific GNI per capita that would have been contributed by the individual over their remaining years of life.

In _daedalus_, the country- and age-specifc VSL values can be accessed from the package data object `daedalus::life_value`.
These values are also automatically accessed and included when a `<daedalus_country>` object is created, under the name `vsl`.

```{r vsl_example}
# load library for examples
library(daedalus)

# create a <daedalus_country>
x <- daedalus_country("China")

# access VSL values using either a helper function or
# the subsetting operator
get_data(x, "vsl")

x$vsl
```
