---
title: "Modelling multiple contact settings"
bibliography: resources/references.json
link-citations: true
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Modelling multiple contact settings}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This vignette shows how to pass community contact matrices disaggregated into multiple settings to the Daedalus model.

The idea is --- in future --- to be able to apply closures or any arbitrary scaling to each setting separately, rather than uniformly scaling contacts across settings.
An example of setting-specific scaling is scaling school contacts for school-age children, while leaving contacts in other settings intact.

**Note that** While the scaffolding for such scaling functionality is in place on the C++ side, this functionality is still being developed on the R side.

## Basic example using POLYMOD

This example using POLYMOD data [@mossong2008] shows how to assign contacts from two different settings to a `<daedalus_country>` object.

Once this assignment is done, helper functions such as `prepare_parameters()` take care of expanding and scaling the contact matrices when the country object is passed to `daedalus()`.

**Note that** running `daedalus("XYZ", ...)` cannot be used to model multiple settings, and the `country` argument must be a `<daedalus_country>` if multiple settings are needed.

```{r setup}
# load package
library(daedalus)
library(socialmixr) # for POLYMOD data
```

By default a country has a single setting, the 'total' community contacts.
This is displayed in the country object when printed to screen.

```{r default_country}
daedalus_country("GBR")
```

We can get POLYMOD data [@mossong2008] for the U.K. from the [R package _socialmixr_](https://cran.r-project.org/package=socialmixr).
Here in this basic example, we consider only two settings, 'home' and 'school'.

```{r get_polymod}
# load polymod data and get home and school contacts
data(polymod)

# get matrices with Daedalus age bins
cm_uk_home <- contact_matrix(
  polymod,
  "GB",
  age_limits = c(0, 5, 18, 65),
  filter = list(cnt_home = 1)
)$matrix

cm_uk_school <- contact_matrix(
  polymod,
  "GB",
  age_limits = c(0, 5, 18, 65),
  filter = list(cnt_school = 1)
)$matrix
```

Create a new country object, passing the new settings to the `contact_matrix` argument as a named list.
Having names helps when examining the object later.

```{r multi_setting}
x <- daedalus_country(
  "GB",
  contact_matrix = list(
    home = cm_uk_home,
    school = cm_uk_school
  )
)

# print x to see output
x
```

Pass the created country to [daedalus()] as usual to run the epi-econ model taking into account contacts in each setting.

```{r run_daedalus}
daedalus(x, "sars_cov_1", time_end = 100)
```

## Future developments

A number of developments to this feature are being considered.
These include:

1. Disaggregating workplace contacts into settings, so that scaling can be applied by setting rather than by economic sector, as this is much easier to work with from a policy perspective (e.g. closing pubs _versus_ scaling hospitality contacts by some amount to represent pub closures).

2. Applying age-specific setting scaling, to represent different policy recommendations or outcomes by age in different settings.

3. Applying different behavioural modifiers to each setting to simulate endogenous risk-perception related modification in contacts across settings.

## References
